---
title: "Examples of usage"
author: "Pablo Rodríguez-Sánchez"
date: "`r Sys.Date()`"
output: 
  pdf_document:
  html_document: default
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, echo = FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(consRvative)
library(latticeExtra)
```

# Installation
To install the package, please type

```
devtools::install_github("PabRod/consRvative", ref = "develop")
```

in your _R_ console.

# One dimensional examples

## Allee effect
A single-species population dynamics model with Allee effect is governed by the following differential equation:

$$
\frac{dN}{dt} = r N \left( \frac{N}{A} - 1 \right) \left( 1 - \frac{N}{K} \right)
$$

It is easy to see that this differential equation has three equilibrium points, $N = 0$, $N = K$ and $N = A$, being all of them stable but the latter one, which is unstable. We'll use the parameters $r = 1$, $A = 0.5$ and $K = 1$.

```{r Allee-def}
r <- 1
A <- 0.5
K <- 1

f <- function(x) { r * x * (x/A - 1) * (1 - x/K) }
```

We can use our method `approxPot1D` to approximate the potential function at a set of points. First, we have to create the points. 

```{r Allee-points}
xs <- seq(0, 1.25, by = 0.01)
```

and then pass them to our algorithm:

```{r Allee-algorithm, cache = TRUE}
Vs <- approxPot1D(f, xs)
```

By plotting the result, we clearly see that the two stable equilibria appear at $N = 0$ and $N = K = 1$, and the unstable one at $N = A = 0.5$, as we expected.

```{r Allee-plot}
plot(xs, Vs, 
     type = 'l', xlab = 'N', ylab = 'V')
```

# Two dimensional examples

## Four well potential
In this section we'll deal with the two-dimensional differential equation given by:

$$
\begin{cases}
\frac{dx}{dt} = f(x,y) = -x(x^2-1) \\
\frac{dy}{dt} = g(x,y) = -y(y^2-1)
\end{cases}
$$

This is a gradient system, that can be derived from the potential:

$$
V(x,y) = \frac{x^2}{4}(x^2 - 2) + \frac{y^2}{4}(y^2 - 2) + V_0
$$

It is important to remember that we'll code the function's input as a vector:

```{r Four-def}
f <- function(x) {c(-x[1]*(x[1]^2 - 1), 
                    -x[2]*(x[2]^2 - 1))}
```

Our region of interest is now two-dimensional. We need, thus, two vectors to create our grid of points:

```{r Four-points}
xs <- seq(-1.5, 1.5, by = 0.05)
ys <- seq(-1.5, 1.5, by = 0.05)
```

Now we are ready to apply `approxPot2D`:

```{r Four-algorithm, cache = TRUE}
result <- approxPot2D(f, xs, ys)
```

`result` is a list that contains two fields:

* `result$V` contains the estimated values of the potentials at each grid point
* `result$err` contains the estimated error at each grid point

```{r Four-plot}
filled.contour(xs, ys, result$V, 
        color=gray.colors, asp = 1,
        main = 'Approximate potential', xlab = 'x', ylab = 'y')
```

```{r Four-plot-lattice, echo = FALSE, eval = FALSE}
data <- expand.grid(xs, ys)
data$Z <- c(result$V)
names(data) <- c('X', 'Y', 'Z')

levelplot(Z  ~ X*Y, data = data, 
          contour = TRUE, col.regions = gray.colors, asp = 1,
          xlab = 'x', ylab = 'y', main = 'Approximate potential')
```

Provided our example is a gradient system, we expect our approximation error to be zero everywhere.

```{r Four-error}
max(result$err)
```

## Simple regulatory gene network
A bistable network model can be described by a set of equations of the form:

$$
\begin{cases}
  \frac{dx}{dt} = f(x,y) = b_x - r_x x + \frac{a_x}{k_x + y^n} \\
  \frac{dy}{dt} = g(x,y) = b_y - r_y y + \frac{a_y}{k_y + x^n}
\end{cases}
$$
We can code it in vector form:

```{r Wadd-def}
# Parameters
bx <- 0.2
ax <- 0.125
kx <- 0.0625
rx <- 1

by <- 0.05
ay <- 0.1094
ky <- 0.0625
ry <- 1

n <- 4

# Dynamics
f <- function(x) {c(bx - rx*x[1] + ax/(kx + x[2]^n), 
                    by - ry*x[2] + ay/(ky + x[1]^n))}
```

This set of equations is, in general, not gradient (because $\frac{\partial f}{\partial y} \neq \frac{\partial g}{\partial x}$). Anyways, we can use the method `approxPot2D` to compute the approximate potential.

First, we need to define our region of interest:

```{r Wadd-points}
xs <- seq(0, 4, by = 0.05)
ys <- seq(0, 4, by = 0.05)
```

And then we are ready to apply our algorithm:

```{r Wadd-algorithm, cache = TRUE}
result <- approxPot2D(f, xs, ys)
```

The resulting approximate potential looks like:

```{r Wadd-plot}
filled.contour(xs, ys, result$V, 
        color = colorRamps::matlab.like, asp = 1,
        main = 'Approximate potential', xlab = 'x', ylab = 'y')

```

Being this not a gradient system it is advisable to plot the estimated error. The areas in green represent small approximation error, so the potential can be safely used in those regions.

```{r Wadd-err}
filled.contour(xs, ys, result$err, 
        color = colorRamps::green2red, asp = 1,
        main = 'Approximation error', xlab = 'x', ylab = 'y')
```

